\section{Introduction}

This report constitutes the final assessment for the Noise and Vibrations course and addresses the vibratory behaviour of a \textbf{sounding rocket} by means of numerical simulation. The geometry analysed is a vertical-launch vehicle developed by the rocketry group of the Universidad de Le\'on, featuring a nose cone, a cylindrical body and fins at the aft end. The slender, elongated configuration of the rocket makes it particularly suitable for a vibration study, since it leads to well-defined bending and torsional modes, and its structural response is directly relevant to flight loads, aerodynamic excitation and potential resonance phenomena during the ascent phase.

The work comprises two analysis blocks. A \textbf{modal analysis} is performed to determine the natural frequencies and mode shapes of the rocket in the absence of external excitation. A \textbf{harmonic analysis} is then carried out to study the structural response to a frequency-dependent sinusoidal excitation, including a comparison of the effect of three different damping values ($\zeta = 1\%$, $\zeta = 5\%$ and $\zeta = 10\%$).

The methodology followed starts with the selection and preparation of the rocket geometry in ANSYS Workbench, then proceeds through the definition of the material (EPOXY E-GLASS UD, consistent with the glass fibre construction of the actual rocket), the generation of a convergent mesh, the application of boundary conditions consistent with the physical problem (a fixed support at the base of the rocket, representing its attachment to the launch pad), and the execution and interpretation of both the modal and harmonic analyses. Each modelling decision is accompanied by the corresponding technical justification.

The goal of this report is not merely the presentation of numerical results, but their \textbf{technical interpretation}: explaining what has been done, why it has been done, what assumptions have been adopted, and what the obtained results mean from a structural and aerospace engineering point of view.
